SDSS-IV MaNGA: Mass Independent Star Formation Enhancement in Close Galaxy Pairs

### 2020-08-04 maintex.tex
next draft

\begin{abstract}
We study the spatial profile of the star formation rate (SFR) in a sample of {\it 160} interacting galaxies within the SDSS-IV MaNGA survey. We find that the specific star formation rate (sSFR) is centrally (e.g. within a single effective radius) enhanced in interacting galaxies by $\sim$0.25 $\pm$ 0.1 but that there is no enhancement or suppression in their disks. We split the radial profile into five separate stellar mass bins from \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass of the paired galaxies. Within one effective radius, the sSFR enhancement in each of the mass bins falls within $\sim$0.1 dex of the median profile for the whole pair sample. We then use a second method to compare paired galaxies and control galaxies where each paired galaxies is matched to a set of 20 control galaxies based on stellar mass and redshift. The sSFR enhancements found are still mass independent, reinforcing the idea that low mass galaxies experience the same level of enhancement to their star formation rate as high mass galaxies. Since the merger induced sSFR enhancement is constant over stellar mass, low mass galaxies will experience a higher fractional mass growth rate of 1$-$7\%, assuming a constant burst merger induced star formation over a 1 Gyr period. This indicates that merger induced star formation has a significant impact on the mass evolution of low mass galaxies.

@@

After the star formation cuts, we require that all galaxies are in the Primary or Secondary MaNGA subsamples \citep{Wake:2017}. The resulting sample cover a stellar mass range of \logm\ $=$ 9.0$-$11.5.

@@

In our previous work, \citet{Fu:2018}, using the 2168 unique galaxies form DR14 \citep{Abolfathi:2018} we showed that $\sim$5.7\% of the MaNGA observations have at least one companion galaxy within the field of view of the IFU. In this work we use a refined pair identification process which is also expanded for the most recent internal data release, MPL-8.

We use SDSS photometric objects to create a catalog of objects within the fields of view of the MaNGA IFUs. We extract their spectra from the MaNGA datacubes through a 1\arcsec\ aperture and fit the spectra with {\sc SPFIT} assuming the MaNGA target's redshift. The spectra is then sorted into different categories; ``good" galaxy spectra, broad-line AGN (active galactic nuclei), foreground star, foreground/background galaxies, or poor S/N objects. The ``good" galaxy spectra are the objects whose spectra is well modeled by {\sc SPFIT} at the target galaxy's redshift, whether it is the target galaxy itself or a nearby companion galaxy. This means that the companion galaxy can be within $\pm$3000 km s$^{-1}$ of the MaNGA target. We found 6573 ``good" objects, 57 broad-line AGN, 836 foreground stars, 319 foreground/background galaxies, and 1546 objects with poor S/N.

From the 6573 galaxies with good spectra, 404 of the MaNGA IFUs have multiple objects within the IFU. We further restrict the sample by setting a relative velocity cut of $\Delta$v $<$ 500 km s$^{-1}$ to remove projected companions and we also require that the galaxies are classified as star forming by the CMD diagram. These requirements reduce our sample down to 124 star forming galaxies (SFGs).

@@

From the NSA catalog's 641,409 galaxies, we find 361 galaxies which are paired to MaNGA targets. MaNGA targets which have both an inside-IFU and an outside-IFU pair are left to the inside-IFU sample. After restricting the sample to SFGs, we have 127 MaNGA targets with paired galaxies outside of the IFU.

\subsection{Control Sample}\label{sec:control}
We build a sample of isolated control galaxies from the remaining MaNGA galaxies by selecting the MaNGA target galaxies which have no spectroscopic companions within r$_p$ $<$ 50 kpc and $\Delta$v $<$ 500 km s$^{-1}$ either inside or outside of the IFU. This gives us a set of 1891 SFGs and 1528 SFGs which are covered by both \citet{Simard:2011} and \citet{Mendel:2014}.

@@

To control for stellar mass, we set a 0.1 dex stellar mass limit between a paired galaxy and the potential control galaxies. To restrict the redshift between the control and paired galaxies, we select control galaxies which are drawn from the same MaNGA subsample (e.g. Primary or Secondary) as the paired galaxy. The MaNGA sample has a tight ``banana" shaped distribution in stellar mass and redshift. By setting a 0.1 dex stellar mass cut and requiring that the galaxies be in the same MaNGA subsample, we effectively set a z $\le$ $\sim$0.025 redshift cut. We note that the MaNGA galaxies with stellar masses above \logm\ $\ge$ 11.0 have a wide redshift distribution, our galaxy samples are restricted to a stellar mass range between \logm\ $=$ 9.0$-$11.0.

@@

The control galaxies have their radial profiles of sSFR made with both the r-band elliptical Petrosian apertures from the NSA catalog and the r-band elliptical S\'ersic apertures from \citet{Simard:2011}. We use the geometries and stellar masses from \citet{Simard:2011} and \citet{Mendel:2014} when the controls are being matched to pairs in the inside-IFU sample and we use the geometries and stellar masses from the NSA catalog when the controls are being matched to pairs in the outside-IFU sample.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}
%\centering
%\includegraphics[width=\linewidth]{fig/bpt_spax.pdf}
%\caption[]{The 2-D histogram showing the BPT classification \citep{Baldwin:1981} of individual spaxels within our star-forming control and pair samples. The color of each bin represents the number density of spaxels, normalized to one. The solid black line represents the theoretically determined maximum starburst line of \citet{Kewley:2001} and the dashed black line is the analytical \citet{Kauffmann:2003} line.}
%\label{fig:bpt}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%% SPFIT
We use our own {\sc IDL} based spectral fitting code, {\sc SPFIT}, to model the spectra from the MaNGA datacubes. {\sc SPFIT} simultaneously fits emission lines and the stellar continuum with the Levenberg-Marquardt nonlinear least-squares minimization algorithm \citep{Fu:2018}. Emission lines are parameterized as a Gauss-Hermite series and the stellar continuum is the sum of simple stellar populations (SSPs) convolved with the line-of-sight velocity distribution (LOSVD). The algorithm provides us with measurements of emission line fluxes, equivalent widths, velocities, and velocity dispersions of 19 emission lines and measurements of stellar masses, ages, [Fe/H], and the kinematics of the stellar populations.

%%% Reddening
We correct the emission lines for reddening using the extinction curve from \citet{Cardelli:1989} with updated coefficients from \citet{ODonnell:1994}. The extinction in parameterized as R$_V$ $\equiv$ A$_V$/E(B$-$V) $=$ 3.1, where we estimate the value of the V-band extinction, A$_V$, by comparing the H$\alpha$/H$\beta$ ratio to the expected value of 2.85 for case-B recombination. The extinction correction is applied to all of the extracted emission lines.

@@

We measure the SFR from the H$\alpha$ luminosity, L$_{H\alpha}$, calculated from the H$\alpha$ flux.  We use the SFR formula, Equation \ref{eq:sfr}, from \citet{Murphy:2011} which uses a Kroupa IMF, Solar metallicity, a constant SFR at an age of 100 Myr, and Case-B recombination.

\begin{equation}\label{eq:sfr}
\frac{\rm{SFR}}{M_{\odot} \, \rm{yr^{-1}}} = \frac{L_{H\alpha}}{1.86 \times 10^{41}\, \rm{erg\,s}^{-1}}
\end{equation}

@@

This will allow us to compare the SFR in high mass regions, like the centers of the galaxies, to the SFR lower mass regions, like the disk of the galaxies.

@@

It has also been shown that hot low mass evolved stars (HOLMES) in retired galaxies can mimic HII regions and AGN on the BPT diagram \citep{Stasinska:2008}. We apply an H$\alpha$ equivalent width (EW) cut of \ewha\ $>$ 3\AA\ to mask retired spaxels \citet{Cid-Fernandes:2011}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Geometry}\label{sec:radial}

@@

We characterize the geometry of the inside-IFU galaxies with the r-band elliptical S\'ersic apertures from \citet{Simard:2011} and we characterize the total stellar mass of the galaxies with the stellar masses from \citet{Mendel:2014}. We use \citet{Simard:2011} instead of the apertures from the NSA catalog, which the MaNGA sample is drawn from, because the apertures in the NSA catalog often fail to fit the MaNGA galaxies when there is another nearby, bright object.

The \citet{Simard:2011} and \citet{Mendel:2014} catalogs do not completely cover the MaNGA sample. Out of the 124 SFGs in the inside-IFU sample, 33 of the SFGs are covered by both \citet{Simard:2011} and \citet{Mendel:2014}.

We use the r-band elliptical Petrosian apertures from the NSA catalog to describe the geometry of the galaxies in the outside-IFU sample. For cases where the identified galaxy pair is outside of the MaNGA IFU, the NSA catalog's algorithm is able to fit the MaNGA target's geometry well. We also use the stellar masses calculated from the r-band elliptical Petrosian apertures in the NSA catalog for the total stellar mass of the galaxies. \\

In order to spatially characterize the star formation in the paired galaxies we build radial profiles of sSFR. As mentioned in Section \ref{sec:data} we use the r-band elliptical Petrosian apertures from the NSA catalog and the r-band S\'ersic apertures from \citet{Simard:2011} to define the geometries of the galaxies. We calculate the inclination angle, {\it i}, of the galaxies using the major-to-minor axis ratios from the elliptical apertures;

@@

The inclination angle, along with the galaxy's position angle, is used to deproject the geometries of the galaxies. The distance to each of the spaxels from the center of the MaNGA target is calculated using the 50\% half light radius to scale the size of the galaxy. The r-band elliptical S\'ersic apertures from \citet{Simard:2011} are used for the inside-IFU sample and their controls and the r-band elliptical Petrosian apertures from the NSA catalog are used for the outside-IFU sample and their controls.

The spaxels are binned into radius increments of 0.1 R/\reff\ from 0.0$-$2.5 R/\reff. Within each radius bin we take the median of the specific star formation rate. The profiles' errors are the standard error of the mean of the data within each radius bin. We create one of these individual averaged radial profiles for every MaNGA galaxy.

We use our catalog of photometric objects and the \citet{Simard:2011} catalog to mask foregournd/background objects which may contaminate the data. Objects in our photometric catalog are masked out with a 2\arcsec\ circular aperture and the galaxies from \citet{Simard:2011} are masked out with a 2.0 R/\reff\ elliptical aperture, using the r-band S\'ersic apertures from the catalog.

@@

\subsection{Mass-Binned Difference Profiles}\label{sec:mass-bin}

In the first method, paired and control galaxies are grouped into four evenly spaced stellar mass bins between \logm\ $=$ 9.0$-$11.0. Within each stellar mass bin, we create a median profile of log(sSFR) as a function of galactocentric radius for both paired and control galaxies. The control profiles are constructed using all available control galaxies within each mass bin. The error associated with the profile is the standard error of mean of the data at each radius bin. We show these ``stacked" profiles for control galaxies in the left hand panel of Figure \ref{fig:ssfr_prof} and the stacked profiles for paired galaxies in the middle panel of Figure \ref{fig:ssfr_prof}.

Once the stacked profiles are made, we take the difference between the stacked profiles of the paired galaxies and the control galaxies, pair - control, in log space (this means that the profiles really represents a ratio between the pairs and controls in linear space). This gives us difference profile, $\Delta$log(sSFR), which shows us where the sSFR is enhanced or suppressed (shown in the right hand panel of Figure \ref{fig:ssfr_prof}). We will discuss the results of the difference profiles later in Section \ref{sec:results}.

\subsection{Tailored-Controls Difference Profiles}\label{sec:tailored}

@@

The difference profiles at this stage have now been controlled for stellar mass, redshift, and galaxy size. We can freely stack these profiles to study the global properties of the merging galaxies. We take all of the $\Delta$log(sSFR) profiles and take the median value within each radius bin to create a stacked profile of the individual difference profiles. We split the individual difference profiles into four evenly spaced stellar mass bins between \logm\ $=$ 9.0$-$11.0 and stack the difference profiles within each mass bin. This gives us four difference profiles covering four different stellar mass ranges shown in the left hand panel of Figure \ref{fig:ssfr_mass}. The results of the profiles will discussed in detail in Section \ref{sec:results}.

To study how the star formation in the centers of paired galaxies are affected by merger events, we extract the $\Delta$log(sSFR) from the individual difference profiles. We extract the value of $\Delta$log(sSFR) from the inner 0.5 R/\reff\ of the paired galaxies and take the mean value of the $\Delta$log(sSFR) within the aperture. We depict the central $\Delta$log(sSFR) as a function of stellar mass in the right hand panel of Figure \ref{fig:ssfr_mass}.

@@

\subsection{Stellar Mass}

In Figures \ref{fig:ssfr_prof} and \ref{fig:ssfr_mass}, we show the profile of $\Delta$log(sSFR) as a function of galactocentric radius between two different methods of comparing paired galaxies to control galaxies. In the first method (Section \ref{sec:mass-bin}) where pair and control galaxies are compared within different stellar mass bins. The difference profiles from this method show a central enhancement to the log(sSFR) of $\sim$0.3 $\pm$ 0.1 dex. The enhancement linearly falls to zero at 1.0 R/\reff\ after which the disks of the paired galaxies have zero enhancement/suppression to their sSFR.

The $\Delta$log(sSFR) profile following the second method shows the same central sSFR enhancement albeit with smaller variances. Within the inner 1.0 R/\reff, there is a variance between the profile of 0.1 dex. Beyond 1.0 R/\reff, the variance increases 0.2 dex. This indicates that the second method of pair$-$ control comparison is superior to the first method.

The profiles of either method show no dependence on the stellar mass of the galaxies.  This effect is better seen in the right hand panel of Figure \ref{fig:ssfr_mass} where the nuclear $\Delta$log(sSFR) of the paired galaxies fall within 0.1 dex of the mean nuclear sSFR enhancement. Low mass galaxies have their star formation enhanced at the same level as high mass galaxies. The similar level of $\Delta$log(sSFR) in high and low mass galaxies will mean that the fractional mass growth between high and low mass galaxies will be different.

@@

To estimate the enhanced mass growth rate due to merger induced star formation, we will assume a constant star formation rate over a given dynamic timescale, $\tau_{dyn}$. A specific star formation rate is already a representation of the fractional mass growth every year, so the merger induced mass growth is simply;

\begin{equation}
\frac{\Delta M_{merger}}{M_{total}} = \left(sSFR_{pair} - sSFR_{control}\right) * \tau_{dyn}
\end{equation}

For this estimation we use a timescale of 1 Gyr, which is the order of the typical timescale between the passage of the first and second pericenters in hydrodynamical simulations \citep{Boylan-Kolchin:2008}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec:disc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=3in]{fig/mass_gain.pdf}
\caption[The fractional mass gain due to merger induced star formation.]{The fractional mass growth due to merger induced star formation over a dynamical timescale, $\tau$ $=$ 1 Gyr. The highlighted region represents the propagated standard errors of the mean of the stacked log(sSFR) profiles. }
\label{fig:mass_gain_sum}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In this work, we primarily focused on how the $\Delta$log(sSFR) in paired galaxies behaves as a function of stellar mass. We also explored how the $\Delta$log(sSFR) behaves as a function of the mass ratio between the paired galaxies and the projected separation between the paired galaxies. While we see hints that the $\Delta$log(sSFR) is affected by the mass ratio and projected separation, these results are less definitive than the $\Delta$log(sSFR) profile's lack of dependence on stellar mass.

\subsection{Mass Ratio}
While the total stellar mass between the galaxies has no effect on the merger induced star formation, the mass ratio between the paired galaxies does. The mass ratio here is calculated as the difference between stellar masses contained within the 50\% half light radius,

@@

\subsection{Projected Separation}\label{sec:sep}

@@

% funding
J.S. and H.F. acknowledge support from the National Science Foundation (NSF) grant AST-1614326 and University of Iowa funds. Funding for the Sloan Digital Sky Survey IV has been provided by the Alfred P. Sloan Foundation, the U.S. Department of Energy Office of Science, and the Participating Institutions. SDSS acknowledges support and resources from the Center for High-Performance Computing at the University of Utah. The SDSS web site is www.sdss.org.

>>>>

\begin{abstract}
We study the spatial profile of the star formation rate (SFR) in a sample of {\it 220} interacting galaxies within the SDSS-IV MaNGA survey. We find that the specific star formation rate (sSFR) is centrally (e.g. within a single effective radius) enhanced in interacting galaxies by $\sim$0.25 $\pm$ 0.1 but that there is no enhancement or suppression in their disks. We split the radial profile into five separate stellar mass bins from \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass of the paired galaxies. Within one effective radius, the sSFR enhancement in each of the mass bins falls within $\sim$0.1 dex of the median profile for the whole pair sample. We then use a second method to compare paired galaxies and control galaxies where each paired galaxies is matched to a set of 20 control galaxies based on stellar mass and redshift. The sSFR enhancements found are still mass independent, reinforcing the idea that low mass galaxies experience the same level of enhancement to their star formation rate as high mass galaxies. Since the merger induced sSFR enhancement is constant over stellar mass, low mass galaxies will experience a higher fractional mass growth rate of 1$-$7\%, assuming a constant burst merger induced star formation over a 1 Gyr period. This indicates that merger induced star formation has a significant impact on the mass evolution of low mass galaxies.

@@

On top of the color selection cuts, we require that all galaxies are in the Primary or Secondary MaNGA subsamples \citep{Wake:2017}. The resulting sample cover a stellar mass range of \logm\ $=$ 9.0$-$11.5.

@@

In our previous work, \citet{Fu:2018}, using the 2168 unique galaxies form DR14 \citep{Abolfathi:2018}, we showed that $\sim$5.7\% of the MaNGA observations have at least one companion galaxy within the field of view of the IFU. In this work we use a refined pair identification process which is also expanded for the most recent internal data release, MPL-8.

We identify potential paired galaxies by overlaying SDSS photometric objects over the MaNGA fields of view. We manually inspect each MaNGA field, removing photometric objects which are over-deblended galaxy fragments and adding any objects missed by the SDSS photometric catalog. At this stage the object catalog includes any foreground stars and foreground/background galaxies along with the potential paired galaxies.

To select paired galaxies out of our object catalog, we inspect the spectra of each object. We use our own {\sc IDL} based spectral fitting code, {\sc SPFIT}, to model the spectra from the MaNGA datacubes. {\sc SPFIT} simultaneously fits emission lines and the stellar continuum with the Levenberg-Marquardt nonlinear least-squares minimization algorithm \citep{Fu:2018}. Emission lines are parameterized as a Gauss-Hermite series and the stellar continuum is the sum of simple stellar populations (SSPs) convolved with the line-of-sight velocity distribution (LOSVD).

The spectra of the identified objects is extracted through a 1\arcsec\ circular aperture and fitted assuming the MaNGA target's redshift and then manually sorted into the following categories; ``good" galaxy spectra, broad-line AGN (active galactic nuclei), foreground star, foreground/background galaxies, or poor S/N objects. The ``good" galaxy spectra are the objects whose spectra is well modeled by {\sc SPFIT} at the target galaxy's redshift, whether it is the target galaxy itself or a nearby companion galaxy. This means that the companion galaxy can be within $\pm$3000 km s$^{-1}$ of the MaNGA target. We found 6573 ``good" objects, 57 broad-line AGN, 836 foreground stars, 319 foreground/background galaxies, and 1546 objects with poor S/N.

From the 6573 galaxies with good spectra, 404 of the MaNGA IFUs have multiple objects within the IFU. We further restrict the sample by setting a relative velocity cut of $\Delta$v $<$ 500 km s$^{-1}$ to remove projected companions. Again, the galaxy sample is also restricted to galaxies with a stellar mass range of \logm\ $=$ 9.0$-$11.5 and galaxies in the Primary or Secondary MaNGA subsamples. We also require that the galaxies are classified as star forming by the CMD diagram. These requirements reduce our sample down to 87 star forming galaxies (SFGs).

@@

From the NSA catalog's 641,409 galaxies, we find 492 galaxies which are paired to MaNGA targets. MaNGA targets which have both an inside-IFU and an outside-IFU pair are left to the inside-IFU sample. After restricting the sample to SFGs, we have 133 MaNGA targets with paired galaxies outside of the IFU.

\subsection{Control Sample}\label{sec:control}

@@

We build a sample of isolated control galaxies from the remaining MaNGA galaxies by selecting the MaNGA target galaxies which have no spectroscopic companions within r$_p$ $<$ 50 kpc and $\Delta$v $<$ 500 km s$^{-1}$ either inside or outside of the IFU. This gives us a set of 1891 SFGs and 1528 SFGs which are covered by both \citet{Simard:2011} and \citet{Mendel:2014}.

We then set two separate methods for comparing the paired galaxies with the control galaxies. In the first method, we simply split the galaxies into separate stellar mass bins. We then compare paired galaxies and control galaxies within each stellar mass bin.

In the second method, we control for both stellar mass and redshift. To do this, we set a 0.1 dex stellar mass limit between a paired galaxy and the potential control galaxies. Then we restrict the redshift between the control and paired galaxies, by selecting control galaxies which are drawn from the same MaNGA subsample (e.g. Primary or Secondary) as the paired galaxy. The MaNGA sample has a tight ``banana" shaped distribution in stellar mass and redshift. By setting a 0.1 dex stellar mass cut and requiring that the galaxies be in the same MaNGA subsample, we effectively set a z $\le$ $\sim$0.025 redshift cut.

@@

While the first method is very basic in comparison to the second method, we will see in Section \ref{sec:results} that the first method is sufficient to see the central enhancement to the star formation rate and the mass independence of the star formation enhancement.

@@

@@

%%% SPFIT
We use {\sc SPFIT} to get measurements of emission line fluxes, equivalent widths, velocities, and velocity dispersions of 19 emission lines and measurements of stellar masses, ages, [Fe/H], and the kinematics of the stellar populations.

%%% Reddening
We correct the emission lines for reddening using the extinction curve from \citet{Cardelli:1989} with updated coefficients from \citet{ODonnell:1994}. The extinction in parameterized as R$_V$ $\equiv$ A$_V$/E(B$-$V) $=$ 3.1, where we estimate the value of the V-band extinction, A$_V$, by comparing the H$\alpha$/H$\beta$ ratio to the expected value of 2.85 for case-B recombination.

\begin{equation}
A_V = 6.78\ {\rm log} \left[(H\alpha/H\beta)/2.85 \right]
\end{equation}

The V-band extinction is then used to calculate the wavelength dependent extinciton A$_{\lambda}$;

\begin{equation}
A_{\lambda}(x) = A_V \left(a(x) + b(x) / R_v \right)
\end{equation}

where $x = 1/\lambda$ and $a(x)$ and $b(x)$ are the coefficients for the extinction curve from \citet{ODonnell:1994}. The extinction corrected flux is then;

\begin{equation}
F'(\lambda) = F(\lambda) * 10^{0.4 A_{\lambda}(x)}
\end{equation}

This extinction correction is applied to all of the extracted emission lines.

@@

We measure the SFR from the H$\alpha$ luminosity, L$_{{\rm H}\alpha}$, calculated from the H$\alpha$ flux.  We use the SFR formula, Equation \ref{eq:sfr}, from \citet{Murphy:2011} which uses a Kroupa IMF, Solar metallicity, a constant SFR at an age of 100 Myr, and Case-B recombination.

\begin{equation}\label{eq:sfr}
\frac{\rm{SFR}}{M_{\odot} \, \rm{yr^{-1}}} = \frac{{\rm L}_{H\alpha}}{1.86 \times 10^{41}\, \rm{erg\,s}^{-1}}
\end{equation}

@@

This will allow us to compare the SFR in high mass regions, like the centers of the galaxies, to the SFR lower mass regions, like the disk of the galaxies. We have also found a strong correlation between sSFR and \ewha and have found the same results using \ewha. This is useful to check as \ewha\ is an observable and independent on {\sc SPFIT}'s measurement of the stellar mass.

@@

It has also been shown that hot low mass evolved stars (HOLMES) in retired galaxies can mimic H\,{\sc ii} regions and AGN on the BPT diagram \citep{Stasinska:2008}. We apply an H$\alpha$ equivalent width (EW) cut of \ewha\ $>$ 3\AA\ to mask retired spaxels \citep{Cid-Fernandes:2011}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Geometry of Galaxies}\label{sec:radial}

@@

In order to spatially characterize the star formation in the paired galaxies we build radial profiles of sSFR. First, the geometry of the galaxies need to be defined.  We use the $r$-band elliptical Petrosian apertures from the NSA catalog and the $r$-band S\'ersic apertures from \citet{Simard:2011} to define the geometries of the galaxies.

The NSA catalog has complete coverage over the MaNGA sample (since MaNGA selects its targets from this catalog); however, it tends to fail to properly fit paired galaxies with close on-sky separations. We found that the apertures from \citet{Simard:2011} work better for these close galaxies, but the catalog does not completely cover the MaNGA sample. We use NSA catalog for the outside-IFU pair sample because they are well separated on the sky and the \citet{Simard:2011} catalog on the inside-IFU sample. In the cases where an Inside-IFU paired galaxy is not covered by \citet{Simard:2011}, we use the ellipse from the NSA catalog.

We fit the geometry with aperture from both catalog when available. When using the first method of pair-control comparison, where pairs and controls are grouped into stellar mass bins, we fit the control galaxies with the NSA apertures. When making the comparison with the second method, paired galaxies fitted with the NSA aperture are compared against controls fitted with the NSA aperture and paired galaxies fitted with the \citet{Simard:2011} apertures are compared against controls fitted with the \citet{Simard:2011} apertures.

We compared the magnitudes extracted from the two apertures for the control galaxies and found that they are in close agreement. This means that the geometries between the two surveys must be in close agreement and the profiles of a galaxy fitted with a NSA aperture can be confidently compared against a galaxy fitted with an aperture from \citet{Simard:2011}.

We calculate the inclination angle, {\it i}, of the galaxies using the major-to-minor axis ratios from the elliptical apertures;

@@

The inclination angle, along with the galaxy's position angle, is used to deproject the geometries of the galaxies. We use the 50\% half light radius, the effective radius (\reff) to scale the sizes of the galaxies. Doing this will allow use the compare galaxies of different sizes against each other.

\subsection{Radial Profiles}

Once the geometry of the galaxies are set, we can build the radial profiles. The spaxels are binned into radius increments of 0.1 R/\reff\ from 0.0$-$2.5 R/\reff. Within each radius bin we take the median of the specific star formation rate. The profiles' errors are the standard error of the mean of the data within each radius bin. We create one of these individual averaged radial profiles for every MaNGA galaxy.

We use our catalog of photometric objects and the \citet{Simard:2011} catalog to mask foreground/background objects which may contaminate the data. Objects in our photometric catalog are masked out with a 2\arcsec\ circular aperture and the galaxies from \citet{Simard:2011} are masked out with a 2.0 R/\reff\ elliptical aperture, using the r-band S\'ersic apertures from the catalog.

\subsection{Star Formation Enhancement}

@@

\subsubsection{Mass-Binned Difference Profiles}\label{sec:mass-bin}

In the first method, paired and control galaxies are grouped into five evenly spaced stellar mass bins between \logm\ $=$ 9.0$-$11.5. Within each stellar mass bin, we create a median profile of log(sSFR) as a function of galactocentric radius for both paired and control galaxies. The control profiles are constructed using all available control galaxies within each mass bin. The error associated with the profile is the standard error of mean of the data at each radius bin. We show these ``stacked" profiles for control galaxies in the left hand panel of Figure \ref{fig:ssfr_prof} and the stacked profiles for paired galaxies in the middle panel of Figure \ref{fig:ssfr_prof}.

Once the stacked profiles are made, we take the difference between the stacked profiles of the paired galaxies and the control galaxies, pair - control, in log space (this means that the profiles really represents a ratio between the pairs and controls in linear space). This gives us difference profile, $\Delta$log(sSFR), which shows us where the sSFR is enhanced or suppressed (shown in the right hand panel of Figure \ref{fig:ssfr_prof}).

This method has the benefit of a high S/N control profile since each mass bin will have hundreds of control galaxies while the median control profiles of the second method will only have 20 controls. On the other hand, this method only controls for a single galaxy parameter, stellar mass. Further the profiles can only be studied as a function of stellar mass so we are unable to study the profiles as a function of projected separation or mass ratio.

We show the mass-binned profiles in Figure \ref{fig:ssfr_prof}. Between the median profiles of the control galaxies and paired galaxies, we can see that the paired profiles are flatted in their centers in comparison to the control profiles. The paired profiles are centrally enhanced by $\sim$0.25 $\pm$ 0.1 dex. This enhancement linearly falls to zero around 1.0 \reff. In the disks the profiles feature a weak suppression to the sSFR of 0.05$-$0.1 dex.

The sSFR enhancement show no dependence on the stellar mass, the profiles of the different mass bins fall within $\sim$0.2 dex of median profile. This means that low mass galaxies are enhanced at the same level as high mass galaxies. It is also notable that while this method of pair-control comparison is not rigorous as the next method, the central star formation enhancement is clearly present.

\subsubsection{Tailored-Controls Difference Profiles}\label{sec:tailored}

@@

The difference profiles at this stage have now been controlled for stellar mass, redshift, and galaxy size. We can freely stack these profiles to study the global properties of the merging galaxies. We take all of the $\Delta$log(sSFR) profiles and take the median value within each radius bin to create a stacked profile of the individual difference profiles. We split the individual difference profiles into five evenly spaced stellar mass bins between \logm\ $=$ 9.0$-$11.5 and stack the difference profiles within each mass bin. This gives us four difference profiles covering four different stellar mass ranges shown in the left hand panel of Figure \ref{fig:ssfr_mass}.

The difference profiles are shown to be centrally enhanced by $\sim$0.30 $\pm$ 0.1 dex. The enhancement falls to zero around 1.0 \reff\ and is slightly suppressed in their disks by $\sim$0.1 dex. The profiles of the pair stellar masses of \logm\ $=$ 9.0$-$9.5 show a strong enhancement in their disks of $\sim$0.1$-$0.2 $\pm$ 0.1 dex.

The profiles are largely consistent with the results of the previous mass binning method. The profiles of this method have less scatter, which may be due to how the paired are more closely matched in stellar mass and redshift. This method also has the added benefit that the profiles can be separated by other parameters instead of just the stellar mass.

We can also inspect the star formation enhancement within individual paired galaxies. We extract the $\Delta$log(sSFR) from the individual difference profiles. We extract the value of $\Delta$log(sSFR) from the inner 0.5 R/\reff\ of the paired galaxies and take the mean value of the $\Delta$log(sSFR) within the aperture. We depict the central $\Delta$log(sSFR) as a function of stellar mass in the right hand panel of Figure \ref{fig:ssfr_mass}. The mass independence can be clearly seen here as the paired galaxies within each stellar mass bin fall within 0.1 dex of the 0.3 dex $\Delta$log(sSFR) median.

@@

@@

To estimate the enhanced mass growth rate due to merger induced star formation, we will assume a constant star formation rate over a given timescale, $\tau$. A specific star formation rate is already a representation of the fractional mass growth every year, so the merger induced mass growth is simply;

\begin{equation}
\frac{\Delta M_{merger}}{M_{total}} = \left({\rm sSFR}_{\rm pair} - {\rm sSFR}_{\rm control}\right) * \tau
\end{equation}

For this estimation we use a timescale of 1 Gyr, which is the order of the typical timescale between the passage of the first and second pericenters in dark-matter simulations \citep{Boylan-Kolchin:2008}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec:disc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=3in]{fig/mass_gain.pdf}
\caption[The fractional mass gain due to merger induced star formation.]{The fractional mass growth due to merger induced star formation over 1 Gyr. The highlighted region represents the propagated standard errors of the mean of the stacked log(sSFR) profiles. }
\label{fig:mass_gain_sum}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We primarily focused on how the $\Delta$log(sSFR) in paired galaxies behaves as a function of stellar mass, but we also explored how the $\Delta$log(sSFR) behaves as a function of the mass ratio between the paired galaxies and the projected separation between the paired galaxies. While we see hints that the $\Delta$log(sSFR) is affected by the mass ratio and projected separation, these results are less definitive than the $\Delta$log(sSFR) profile's lack of dependence on stellar mass.

\subsection{Dependency on Mass Ratio}
While the total stellar mass between the galaxies has no effect on the merger induced star formation, the mass ratio between the paired galaxies does. The mass ratio here is calculated as the difference between stellar masses contained within the 50\% half light radius,

@@

\subsection{Dependency on Projected Separation}\label{sec:sep}

@@

% funding
J.S. and H.F. acknowledge support from the National Science Foundation (NSF) grant AST-1614326. Funding for the Sloan Digital Sky Survey IV has been provided by the Alfred P. Sloan Foundation, the U.S. Department of Energy Office of Science, and the Participating Institutions. SDSS acknowledges support and resources from the Center for High-Performance Computing at the University of Utah. The SDSS web site is www.sdss.org.
